---
title: "Assignment 5 - Meta-analysis of pitch in schizophrenia"
author: "Simon Moeller Nielsen"
date: "09/12/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Setting WD
setwd("C:/Users/slmoni/Documents/Uni/Experimental Methods III/EXPMETH3/Assignment 5")

# Loading packages
library(pacman)
p_load(tidyverse, metafor, lmerTest, ggbeeswarm)

# Loading data
data = read.csv("SR_SCHIZO.csv", stringsAsFactors = FALSE)
```

# Building on the shoulders of giants: meta-analysis

## Questions to be answered

1. What is the current evidence for distinctive patterns of pitch mean and pitch sd in schizophrenia? Report how many papers report quantitative estimates, your method to analyze them, the estimated effect size of the difference (mean effect size and standard error for pitch mean, same for pitch sd) and forest plots representing it. 
```{r 1}
## For fundamental frequency matadata
pitchF0 = escalc('SMD', n1i=SAMPLE_SIZE_SZ, n2i=SAMPLE_SIZE_HC, m1i=PITCH_F0_SZ_M, m2i=PITCH_F0_HC_M, sd1i=PITCH_F0_SZ_SD, sd2i=PITCH_F0_HC_SD, data = data)

# Calculateing effect size and sd with Lmer
pitchF0Model = lmer(yi ~ 1 + (1 | StudyID), weights = 1/vi, data=pitchF0, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(pitchF0Model)

# Now using rma
pitchF0RMA <-rma(yi, vi, data = pitchF0, slab=Article)
pitchF0RMA

# Forest plot
forest(pitchF0RMA)

## For range matadata
pitchSD = escalc('SMD', n1i=SAMPLE_SIZE_SZ, n2i=SAMPLE_SIZE_HC, m1i=PITCH_F0SD_SZ_M, m2i=PITCH_F0SD_HC_M, sd1i=PITCH_F0SD_SZ_SD, sd2i=PITCH_F0SD_HC_SD, data = data)

# Calculateing effect size and sd with Lmer
pitchSDModel = lmer(yi ~ 1 + (1 | StudyID), weights = 1/vi, data=pitchSD, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(pitchSDModel)

# Now using rma
pitchSDRMA <-rma(yi, vi, data = pitchSD, slab=Article)
pitchSDRMA

# Forest plot
forest(pitchSDRMA)
```


2. Do the results match your own analysis from Assignment 3? If you add your results to the meta-analysis, do the estimated effect sizes change? Report the new estimates and the new forest plots.
```{r 2}
# Loading data
pitchData<-read.csv("SimonPitchData.csv")

# Log transforming the data
pitchData$Mean<-log(pitchData$Mean)
pitchData$SD<-log(pitchData$SD)

# Standardize the data to cohen's d
pitchData$Mean<-(pitchData$Mean-mean(pitchData$Mean))/sd(pitchData$Mean)
pitchData$SD<-(pitchData$SD-mean(pitchData$SD))/sd(pitchData$SD)

# Making the models
meanModel = glmer(Diagnosis ~ Mean + (1|Subject) + (1|Triangles), data= pitchData, family="binomial")
SDModel = glmer(Diagnosis ~ SD + (1|Subject) + (1|Triangles), data= pitchData, family="binomial")
summary(meanModel)
summary(SDModel)

# Adding assignment 3 results
data <- rbind(data, rep(NA,length(data)))

data$SAMPLE_SIZE_SZ[56] = 86
data$SAMPLE_SIZE_HC[56] = 86
data$Article[56] = "cogsci_fun_with_malte" #good one
data$StudyID[56] = 49
data$ArticleID[56] = 47
data$Year_publication[56] = 2018
data$Authors[56] = "Simon, Laurits, Osfar, Signe Kløve, and Thea"

pitchData<-read.csv("SimonPitchData.csv")

group_by(pitchData, Diagnosis) %>%
  summarise(mean(na.omit(Mean)))

group_by(pitchData, Diagnosis) %>%
  summarise(sd(na.omit(Mean)))

group_by(pitchData, Diagnosis) %>%
  summarise(mean(na.omit(SD)))

group_by(pitchData, Diagnosis) %>%
  summarise(sd(na.omit(SD)))

data$PITCH_F0_HC_M[56] = 141
data$PITCH_F0_HC_SD[56] = 49.2
data$PITCH_F0_SZ_M[56] = 149
data$PITCH_F0_SZ_SD[56] = 55.6

data$PITCH_F0SD_HC_M[56] = 25.8
data$PITCH_F0SD_HC_SD[56] = 18.7
data$PITCH_F0SD_SZ_M[56] = 22.9
data$PITCH_F0SD_SZ_SD[56] = 33.0

## Runing the mata analysis again
## For fundamental frequency matadata
pitchF0 = escalc('SMD', n1i=SAMPLE_SIZE_SZ, n2i=SAMPLE_SIZE_HC, m1i=PITCH_F0_SZ_M, m2i=PITCH_F0_HC_M, sd1i=PITCH_F0_SZ_SD, sd2i=PITCH_F0_HC_SD, data = data)

# Now using rma
pitchF0RMA <-rma(yi, vi, data = pitchF0, slab=Article)
pitchF0RMA

# Forest plot
forest(pitchF0RMA)

## For range matadata
pitchSD = escalc('SMD', n1i=SAMPLE_SIZE_SZ, n2i=SAMPLE_SIZE_HC, m1i=PITCH_F0SD_SZ_M, m2i=PITCH_F0SD_HC_M, sd1i=PITCH_F0SD_SZ_SD, sd2i=PITCH_F0SD_HC_SD, data = data)

# Now using rma
pitchSDRMA <-rma(yi, vi, data = pitchSD, slab=Article)
pitchSDRMA

# Forest plot
forest(pitchSDRMA)

```
3. Assess the quality of the literature: report and comment on heterogeneity of the studies (tau, I2), on publication bias (funnel plot), and on influential studies.

```{r 2}
# Getting the tau and I2 for the mean model
pitchF0RMA

# Getting the tau and I2 for the SD model
pitchSDRMA

## Funnel plots
# For mean model
funnel(pitchF0RMA, main = "Random-Effects Model: Mean", xlab = "Standardized Mean Difference")

regtest(pitchF0RMA)
ranktest(pitchF0RMA)

# For SD model
funnel(pitchSDRMA, main = "Random-Effects Model: SD", xlab = "Standardized Mean Difference")

regtest(pitchSDRMA)
ranktest(pitchSDRMA)

## Looking at influential studies
# For the mean model
infF0 = influence(pitchF0RMA)
infF0
plot(infF0)
# And for the SD model
infSD = influence(pitchSDRMA)
infSD
plot(infSD)
```

## Tips on the process to follow:

- Download the data on all published articles analyzing pitch in schizophrenia (on gitlab)
- Look through the dataset to find out which columns to use, and if there is any additional information written as comments (real world data is always messy!).
    * Hint: Make sure you read the comments in the columns: `pitch_f0_variability`, `frequency`, `Title`,  `ACOUST_ANA_DESCR`, `DESCRIPTION`, and `COMMENTS`
- Following the procedure in the slides calculate effect size and standard error of the effect size per each study. N.B. we focus on pitch mean and pitch standard deviation.
 . first try using lmer (to connect to what you know of mixed effects models)
 . then use rma() (to get some juicy additional statistics)

- Build a forest plot of the results (forest(model))
 
- Go back to Assignment 3, add your own study to the data table, and re-run meta-analysis. Do the results change?

- Now look at the output of rma() and check tau and I2
